---
title: <font size="5">**SMCVR -- Checks and exclusions**</font> 
author: <br> <font size="4"> Pawel Motyka (Polish Academy of Sciences) </font> <br>  *pmotyka@psych.pan.pl* 
date: <font size="3"> September 2022  </font>
output: html_document
chunk_output_type: console

--- 
&nbsp;
<font size="4">
**List of sections**:

1. Load the required packages and data (CFS & Sensorimotor Mastery tasks) [S1](#S1)
2. CFS - check the data with respect to the preregistered exclusion criteria [S2](#S2)
3. CFS - additional checks [S3](#S3)
4. Sensorimotor Mastery Tasks - check the data with respect to the preregistered exclusion criteria [S4](#S4)

<a name="S1"></a>
&nbsp;

#####**1. Load the required packages and data (CFS & Sensorimotor Mastery tasks)** 

```{r, message = FALSE, warning = FALSE}

# load packages
library(here, warn.conflicts = FALSE, quietly=TRUE)
library(dplyr, warn.conflicts = FALSE, quietly=TRUE)
library(lattice, warn.conflicts = FALSE, quietly=TRUE)
library(afex, warn.conflicts = FALSE, quietly=TRUE)
library(emmeans, warn.conflicts = FALSE, quietly=TRUE)

# set working directory
data_dir <- paste0(here(),"/_data")
setwd(data_dir)
options(Encoding="UTF-8")

# specify ID list
ID_list <- c(1:48)

# read individual data within the loop
data <- NULL
for (ID in ID_list) {
#print(ID)
log <- read.csv(paste('SMCVR_CFS_', ID, '.csv', sep=''), header = TRUE, sep = ",") 
data <- rbind(data, log)
}

# remove training trials
data <- data[data$training == "False",]
rm(log)

# Read preprocessed data (takes a while)
#SMC_moving <- read.table(file = "smcvr_smc_moving.csv")
#SMC_static <- read.table(file = "smcvr_smc_static.csv")

# Optional step: recreate preprocessed data by reading individual data files within the loop (to do this, change F to T)
Read_raw_data = T 

if (Read_raw_data == T) {

ID_list <- c(1:48)
dm <- NULL
ds <- NULL
for (ID in ID_list) {
#print(ID)

log_dm <- read.csv(paste('SMCVR_SMC_moving_', ID, '.csv', sep=''), header = TRUE, sep = ",") 

log_ds <- read.csv(paste('SMCVR_SMC_static_', ID, '.csv', sep=''), header = TRUE, sep = ",") 

dm <- rbind(dm, log_dm)
ds <- rbind(ds, log_ds)
}

# remove training trials
dm <- dm[dm$training == "False",]
ds <- ds[ds$training == "False",]

write.table(dm, file = "smcvr_smc_moving.csv" )
write.table(ds, file = "smcvr_smc_static.csv" ) 
}

# get demographic data
demo <- data %>%
    group_by(ID) %>%  
    slice(1) %>%
    select(ID,age,sex,eye_d)

# composition
length(demo$ID[demo$sex == "Male"])
length(demo$ID[demo$sex == "Female"])

# age
mean(demo$age)
sd(demo$age)
range(demo$age)

# dominant eye
length(demo$ID[demo$eye_d == "Left"])
length(demo$ID[demo$eye_d == "Right"])
rm(demo)

```

<a name="S2"></a>
&nbsp;

#####**2. CFS - check the data with respect to the preregistered exclusion criteria** 

```{r}

### Criterion (a) "More than 45% of trials with no detection in any of the four conditions (i.e., particular objects). Justification: this would indicate extreme perceptual suppression and impinge on the calculation of median detection times for a given participant and condition".

dat = data
cfs1 <- data.frame(ID = integer(0),
                  condition = integer(0),
                  trials_num = integer(0),
                  detected = integer(0),
                  proportion = numeric(0))

for (i in unique(dat$ID)) {
  
ds <- dat[dat$ID == i,]
#print(paste("###ID###", i))

for (c in unique(ds$condition)) {
trials_num <- length(ds$ID[ds$condition == c])
detected <- length(ds$ID[ds$condition == c & ds$detected == "True"])  
proportion <- detected/trials_num

cfs1[nrow(cfs1)+1,] <- c(i,c,trials_num,detected, proportion)
}
}

# summary (detection)
cfs1$proportion <- as.numeric(cfs1$proportion)
mean(cfs1$proportion)
hist(cfs1$proportion, col = "darkgreen", xlab = "Proportion of trials with detected stimulus", main = NULL)


### Criterion (b) "More than 15% of trials with incorrect identification of the reportedly detected objects. Justification: this would suggest that the subject did not pay sufficient attention during the task."

dat = data
cfs2 <- data.frame(ID = integer(0),
                  trials_num = integer(0),
                  incorrectly_identified = integer(0),
                  proportion = numeric(0))

for (i in unique(dat$ID)) {
  
ds <- dat[dat$ID == i,]
#print(paste("###ID###", i))

trials_num <- length(ds$ID)
incorrectly_identified <- length(ds$ID[ds$detected == "True" & ds$accuracy == "False"])  
proportion <- incorrectly_identified/trials_num

cfs2[nrow(cfs2)+1,] <- c(i, trials_num, incorrectly_identified, proportion)
}

# summary (identification errors)
mean(cfs2$proportion)
hist(cfs2$proportion, col = "darkred", xlab = "Proportion of trials with identification errors", main = NULL)


### Criterion (c) "Median detection time smaller than 500 ms (all trials considered).Justification: this would indicate unusually weak or no genuine perceptual suppression."

cfs3 <- data %>% group_by(ID) %>% dplyr::summarize(t_md = median(detection))

# summary (detection times)
mean(cfs3$t_md)
hist(cfs3$t_md, breaks = 15, col = "darkblue", xlab = "Median detection times", main = NULL)

```

<a name="S3"></a>
&nbsp;

#####**3. CFS - additional checks** 

```{r}

### Check the counterbalancing - objects x conditions
pairs1 <- data %>% group_by(ID, condition, obj_type) %>% dplyr::summarize(t_md = median(detection))
pairs2 <- pairs1[, c(1,2,3)]

pairs <- pairs2 %>% tidyr::pivot_wider(names_from = condition, values_from = obj_type)
rm(dat)
rm(ds)

for (o in 1:4) {
print(length(pairs$ID[pairs$Congruent == o]))
print(length(pairs$ID[pairs$Incongruent == o]))
print(length(pairs$ID[pairs$Novel == o]))
print(length(pairs$ID[pairs$Random == o]))
}


### Proportion of different responses (visual summary)
dat = data
cfs <- data.frame(ID = integer(0),
                  undetected = integer(0),
                  incorrect = integer(0),
                  proportion = numeric(0))

for (i in unique(dat$ID)) {
  
ds <- dat[dat$ID == i,]
#print(paste("###ID###", i))

trials_num <- length(ds$ID)
undetected <- length(ds$ID[ds$detected == "False"])  
incorrect <- length(ds$ID[ds$detected == "True" & ds$accuracy == "False"])
proportion <- undetected/trials_num

cfs[nrow(cfs)+1,] <- c(i,undetected, incorrect, proportion)

}

hist(cfs$undetected, breaks = 48, col = "gray20")
hist(cfs$incorrect, breaks = 48, col = "gray20")
hist(cfs2$incorrectly_identified, breaks = 48, col = "gray20")
sum(cfs$undetected)

```

<a name="S4"></a>
&nbsp;

#####**4. Sensorimotor Mastery Tasks - check the data with respect to the preregistered exclusion criteria** 

```{r}

# Rename Sensorimotor Mastery data
dm <- SMC_moving
ds <- SMC_static

### Criterion d) Excessively slow or fast velocity of the manual actions during the sensorimotor mastery tasks defined as the average speed outside the expected range signalled by the display of warnings (i.e., lower than 0.25 degrees/frame or higher than 3 degrees/frame).

# Moving Target Version

dm_speed <- dm %>%
    #For each Class
    group_by(ID) %>%  
    arrange(desc(timestamp)) %>% 
    slice(1)

dm_speed <- select(dm_speed, ID, speed_average)

hist(dm_speed$speed_average, breaks = 20, xlim = c(0.2, 3), col = "darkorange", xlab = "Moving Target - speed of action", main = NULL)

# Static Target Version

ds_speed <- ds %>%
    #For each Class
    group_by(ID) %>%  
    arrange(desc(timestamp)) %>% 
    slice(1)

ds_speed <- select(ds_speed, ID, speed_average)

hist(ds_speed$speed_average, breaks = 20, xlim = c(0.2, 3), col = "darkorange", xlab = "Static Target - speed of action", main = NULL)


### Criterion e) No better performance in the congruent condition as compared to other conditions (taken together) during the sensorimotor mastery tasks. Justification: meeting criteria d) and/or e) would suggest that the participant did not pay sufficient attention during the tasks.

# Moving Target Version

dmm <- dm %>% group_by(ID, block, condition) %>% dplyr::summarize(d_m = mean(distance), d_md = median(distance))

dt <- dmm

dt <- dt %>% group_by(ID, condition) %>% dplyr::summarize(d_md = median(d_md))

dt <- dt %>% tidyr::pivot_wider(names_from = condition, values_from = d_md)

dt$exclusion <- dt$Congruent > mean(c(dt$Incongruent, dt$Novel, dt$Random))

# additional check: better performance in the Novel than the Random condition
dt$novel_learned <- dt$Novel < dt$Random
dt$novel_learned2 <- dt$Random - dt$Novel
#hist(dt$novel_learned2)

# Static Target Version
dss <- ds %>%
    group_by(ID, block, trial) %>%  
    arrange(desc(timestamp)) %>% 
    slice(1)

dss <- dss %>% group_by(ID, condition) %>% dplyr::summarize(h_m = mean(target_num))

dt <- dss %>% tidyr::pivot_wider(names_from = condition, values_from = h_m)

dt$exclusion <- dt$Congruent <= mean(c(dt$Incongruent, dt$Novel, dt$Random))

# additional check: better performance in the Novel than the Random condition
dt$novel_learned <- dt$Novel > dt$Random
dt$novel_learned2 <- dt$Novel - dt$Random
#hist(dt$novel_learned2)

```


