---
title: <font size="5">**SMCVR -- Sensorimotor Mastery Tasks data**</font> 
author: <br> <font size="4"> Pawel Motyka (Polish Academy of Sciences) </font> <br>  *pmotyka@psych.pan.pl* 
date: <font size="3"> September 2022  </font>
output: html_document
chunk_output_type: console

--- 
&nbsp;
<font size="4">
**List of sections**:

1. Load the required packages and data [S1](#S1)
2. Performance under different sensorimotor conditions - Moving Target version [S2](#S2)
3. Performance under different sensorimotor conditions - Static Target version [S3](#S3)
4. Performance for different visual objects (Moving & Static versions) [S4](#S4)
5. Block-by-block changes in performance (Moving & Static versions) [S5](#S5)
6. Relationship between performance in Moving and Static Target versions [S6](#S6)
7. Relationship between performance and speed of actions [S7](#S7)
8. Merging CFS and Sensorimotor Mastery data [S8](#S8)

<a name="S1"></a>
&nbsp;

#####**1. Load the required packages and data** 

```{r, message = FALSE, warning = FALSE}

# load packages
library(here, warn.conflicts = FALSE, quietly=TRUE)
library(dplyr, warn.conflicts = FALSE, quietly=TRUE)
library(lattice, warn.conflicts = FALSE, quietly=TRUE)
library(tidyr, warn.conflicts = FALSE, quietly=TRUE)
library(afex, warn.conflicts = FALSE, quietly=TRUE)
library(emmeans, warn.conflicts = FALSE, quietly=TRUE)
library(ggpubr, warn.conflicts = FALSE, quietly=TRUE)

# set working directory
data_dir <- paste0(here(),"/_data")
setwd(data_dir)

# read preprocessed data (takes a while)
dm <- read.table(file = "smcvr_smc_moving.csv" )
ds <- read.table(file = "smcvr_smc_static.csv" )

# Optional step: recreate preprocessed data by reading individual data files within the loop (to do this, change F to T)
Read_raw_data = F 

if (Read_raw_data == T) {

  ### Recreate from raw data
# specify ID list
ID_list <- c(1:48)
dm <- NULL
ds <- NULL
for (ID in ID_list) {
print(ID)

# read individual dataframes within the loop
log_dm <- read.csv(paste('SMCVR_SMC_moving_', ID, '.csv', sep=''), header = TRUE, sep = ",")
log_ds <- read.csv(paste('SMCVR_SMC_static_', ID, '.csv', sep=''), header = TRUE, sep = ",")

dm <- rbind(dm, log_dm)
ds <- rbind(ds, log_ds)
}

# remove training trials
dm <- dm[dm$training == "False",]
ds <- ds[ds$training == "False",]
#write.table(dm, file = "smcvr_smc_moving.csv" )
#write.table(ds, file = "smcvr_smc_static.csv" ) 
}


# define colours to be used
con = rgb(.28,.13,.45, alpha = 0.65)
opp = rgb(.18,.44,.56, alpha = 0.8)
nov = rgb(.136,.553,.425, alpha = 0.8)
ran = rgb(.70,.64,.15, alpha = 0.8)
dark = rgb(.1,.1,.1, alpha = 0.9)

conL = rgb(.28,.13,.45, alpha = 0.4)
oppL = rgb(.18,.44,.56, alpha = 0.5)
novL = rgb(.136,.553,.425, alpha = 0.5)
ranL = rgb(.70,.64,.15, alpha = 0.5)

mycol <- rgb(100, 100, 100, max = 255, alpha = 80)
mcol <- rgb(100, 100, 100, max = 255, alpha = 120)
  

# read normalised CFS data (to be merged into a new data file)
cfs <- read.table(file = "smcvr_cfs_normalised.csv")
cfs_ID <- read.table(file = "smcvr_cfs_ID_normalised.csv")

```

<a name="S2"></a>
&nbsp;

#####**2. Performance under different sensorimotor conditions (Moving Target version)** 

```{r}

### prepare data
dt <- dm %>% group_by(ID, condition) %>% dplyr::summarize(d_md = median(distance), d_m = mean(distance))
dt <- select(dt, ID, condition,d_md)
dt$d_md <- dt$d_md * 100 # cm 
dp <- tidyr::pivot_wider(dt, names_from = condition, values_from = d_md)


### analyze differences
m <- aov_ez(id = "ID", dv = "d_md", within = "condition", data = dt, return = "nice")
m # "return = nice" -> only to see GES values

m <- aov_ez(id = "ID", dv = "d_md", within = "condition", data = dt)
summary(m) # degrees of freedom: levels - 1,  (subjects - 1) * (levels - 1)

### post-hoc comparisons
em <- emmeans(m, ~ condition)
em
em_pairs <- pairs(em, adjust = "Bonferroni") 
em_pairs


### visualise results
cond1 <- dp$Congruent
cond2 <- dp$Incongruent
cond3 <- dp$Novel
cond4 <- dp$Random

# greater margins, optionally: subplots mfrow
par(mar=c(1.1,5.1,1.1,1.1),mfrow=c(1,1))

# horizontal jitter across each condition
x <- jitter(rep(0,length(cond2)), factor = 0.4) # narrower

# plotting area
plot(x, cond2, ylim=c(0,max(c(cond1,cond2,cond3,cond4))*1.1), xlim=c(-.25,.25), type="n", ylab="Distance", xlab=NA, xaxt="n", cex.lab=2, cex.axis=1.5)
 
# Plot individual lines
for (i in 1:length(cond2)){
	points( x[i]+c(-0.18,-0.06,0.06,0.18), cbind(cond1,cond2,cond3,cond4)[i,], col=c(con,opp,nov,ran),bg=c(conL,oppL,novL,ranL), pch=21 )
	lines( x[i]+c(-0.18,-0.06,0.06,0.18), cbind(cond1,cond2,cond3,cond4)[i,], col=rgb(.5,.5,.5,alpha=.5) )
}

# calculate means
means <- colMeans(cbind(cond1,cond2,cond3,cond4))

# plot means
lines( c(-.192,-.168), rep(means[1],2), lwd=3.5, col = dark, lty = 1)
lines( c(-.072,-.048), rep(means[2],2), lwd=3.5, col = dark, lty = 1)
lines( c(.048,.072), rep(means[3],2), lwd=3.5, col = dark, lty = 1)
lines( c(.168,.192), rep(means[4],2), lwd=3.5,col = dark, lty = 1)

# optional dots
#points(-.18, means[1], cex = 1.2, pch = 21, col = "black", bg = dark)
#points(-.06, means[2], cex = 1.2, pch = 21, col = "black", bg = dark)
#points(.06, means[3], cex = 1.2, pch = 21, col = "black",  bg = dark)
#points(.18, means[4], cex = 1.2, pch = 21, col = "black",  bg = dark)

# export 4.5 x 5.6 cm



```

<a name="S3"></a>
&nbsp;

#####**3. Performance under different sensorimotor conditions (Static Target version)** 

```{r}

### prepare data
dss <- ds %>%
    group_by(ID, block, trial) %>%  
    arrange(desc(timestamp)) %>% 
    slice(1) # take the last row from each trial to get the number of hits

dss <- dss %>% group_by(ID, condition) %>% dplyr::summarize(h_m = mean(target_num), h_md = median(target_num))
dt <- dss


## analyze differences
m <- aov_ez(id = "ID", dv = "h_m", within = "condition", data = dt, return = "nice")
m # only to see GES values

m <- aov_ez(id = "ID", dv = "h_m", within = "condition", data = dt)
summary(m) # degrees of freedom: levels - 1,  (subjects - 1) * (levels - 1)

# post-hoc comparisons
em <- emmeans(m, ~ condition)
em
em_pairs <- pairs(em, adjust = "Bonferroni") 
em_pairs


### visualise results
dt <- select(dt, ID, condition,h_m)
dp <- tidyr::pivot_wider(dt, names_from = condition, values_from = h_m)

cond1 <- dp$Congruent
cond2 <- dp$Incongruent
cond3 <- dp$Novel
cond4 <- dp$Random

# greater margins, optionally: subplots mfrow
par(mar=c(4.1,5.1,1.1,1.1),mfrow=c(1,1))

# horizontal jitter across each condition
x <- jitter(rep(0,length(cond2)), factor = 0.4) # narrower

# plotting area
plot(x, cond2, ylim=c(0,max(c(cond1,cond2,cond3,cond4))*1.1), xlim=c(-.25,.25), type="n", ylab="Hits", xlab=NA, xaxt="n", yaxt="n", cex.lab=2, cex.axis=1.5)

y_ticks <- c(0,6,12,18)  
axis(side = 2, at = y_ticks, labels = y_ticks, cex.axis = 1.5)

# Plot individual lines
for (i in 1:length(cond2)){
	points( x[i]+c(-0.18,-0.06,0.06,0.18), cbind(cond1,cond2,cond3,cond4)[i,], col=c(con,opp,nov,ran),bg=c(conL,oppL,novL,ranL), pch=21 )
	lines( x[i]+c(-0.18,-0.06,0.06,0.18), cbind(cond1,cond2,cond3,cond4)[i,], col=rgb(.5,.5,.5,alpha=.5) )
}

# calculate means
means <- colMeans(cbind(cond1,cond2,cond3,cond4))

# plot means
lines( c(-.192,-.168), rep(means[1],2), lwd=3.5, col = dark, lty = 1)
lines( c(-.072,-.048), rep(means[2],2), lwd=3.5, col = dark, lty = 1)
lines( c(.048,.072), rep(means[3],2), lwd=3.5, col = dark, lty = 1)
lines( c(.168,.192), rep(means[4],2), lwd=3.5,col = dark, lty = 1)

# optional dots
#points(-.18, means[1], cex = 1.2, pch = 21, col = "black", bg = dark)
#points(-.06, means[2], cex = 1.2, pch = 21, col = "black", bg = dark)
#points(.06, means[3], cex = 1.2, pch = 21, col = "black",  bg = dark)
#points(.18, means[4], cex = 1.2, pch = 21, col = "black",  bg = dark)

# export 5 x 5.5 cm

```

<a name="S4"></a>
&nbsp;

#####**4. Performance for different visual objects (Moving & Static versions)** 

```{r}

### Moving Target Version
 
## prepare data
dt <- dm %>% group_by(ID, obj_type) %>% dplyr::summarize(d_md = median(distance))
dt <- select(dt, ID, obj_type,d_md)
dt$d_md <- dt$d_md * 100

## Analyze differences
m <- aov_ez(id = "ID", dv = "d_md", within = "obj_type", data = dt, return = "nice")
m # only to see GES values

m <- aov_ez(id = "ID", dv = "d_md", within = "obj_type", data = dt)
summary(m) # degrees of freedom: levels - 1,  (subjects - 1) * (levels - 1)

# post-hoc comparisons
em <- emmeans(m, ~ obj_type)
em
em_pairs <- pairs(em, adjust = "Bonferroni")
em_pairs

# visualise results
dt$obj_type[dt$obj_type == 1] <- "O1"
dt$obj_type[dt$obj_type == 2] <- "O2"
dt$obj_type[dt$obj_type == 3] <- "O3"
dt$obj_type[dt$obj_type == 4] <- "O4"

dp <- tidyr::pivot_wider(dt, names_from = obj_type, values_from = d_md)

cond1 <- dp$O1
cond2 <- dp$O2
cond3 <- dp$O3
cond4 <- dp$O4

# greater margins, optionally: subplots mfrow
par(mar=c(1.1,5.1,1.1,1.1),mfrow=c(1,1))

# horizontal jitter across each condition
x <- jitter(rep(0,length(cond2)), factor = 0.4) # narrower

# plotting area
plot(x, cond2, ylim=c(0,max(c(cond1,cond2,cond3,cond4))*1.1), xlim=c(-.25,.25), type="n", ylab="Distance", xlab=NA, xaxt="n", cex.lab=2, cex.axis=1.5)
 
# Plot individual lines
for (i in 1:length(cond2)){

	points( x[i]+c(-0.18,-0.06,0.06,0.18), cbind(cond1,cond2,cond3,cond4)[i,], col=c(con,opp,nov,ran),bg=c(conL,oppL,novL,ranL), pch=21 )
	lines( x[i]+c(-0.18,-0.06,0.06,0.18), cbind(cond1,cond2,cond3,cond4)[i,], col=rgb(.5,.5,.5,alpha=.5) )
}

# calculate means
means <- colMeans(cbind(cond1,cond2,cond3,cond4))

# plot means
lines( c(-.192,-.168), rep(means[1],2), lwd=3.5, col = dark, lty = 1)
lines( c(-.072,-.048), rep(means[2],2), lwd=3.5, col = dark, lty = 1)
lines( c(.048,.072), rep(means[3],2), lwd=3.5, col = dark, lty = 1)
lines( c(.168,.192), rep(means[4],2), lwd=3.5,col = dark, lty = 1)


### Static Target Version

dss <- ds %>%
    group_by(ID, block, trial) %>%  
    arrange(desc(timestamp)) %>% 
    slice(1)
dss <- dss %>% group_by(ID, obj_type) %>% dplyr::summarize(h_m = mean(target_num))
dt <- dss


## analyze differences
m <- aov_ez(id = "ID", dv = "h_m", within = "obj_type", data = dt, return = "nice")
m # only to see GES values

m <- aov_ez(id = "ID", dv = "h_m", within = "obj_type", data = dt)
summary(m) # degrees of freedom: levels - 1,  (subjects - 1) * (levels - 1)

# post-hoc comparisons
em <- emmeans(m, ~ obj_type)
em
em_pairs <- pairs(em, adjust = "Bonferroni") 
em_pairs


### visualise results
dt$obj_type[dt$obj_type == 1] <- "O1"
dt$obj_type[dt$obj_type == 2] <- "O2"
dt$obj_type[dt$obj_type == 3] <- "O3"
dt$obj_type[dt$obj_type == 4] <- "O4"

dp <- tidyr::pivot_wider(dt, names_from = obj_type, values_from = h_m)

cond1 <- dp$O1
cond2 <- dp$O2
cond3 <- dp$O3
cond4 <- dp$O4

# greater margins, optionally: subplots mfrow
par(mar=c(4.1,5.1,1.1,1.1),mfrow=c(1,1))

# horizontal jitter across each condition
x <- jitter(rep(0,length(cond2)), factor = 0.4) # narrower

# plotting area
plot(x, cond2, ylim=c(0,max(c(cond1,cond2,cond3,cond4))*1.1), xlim=c(-.25,.25), type="n", ylab="Hits", xlab=NA, xaxt="n", yaxt = "n", cex.lab=2, cex.axis=1.5)

y_ticks <- c(0,6,12,18)  
axis(side = 2, at = y_ticks, labels = y_ticks, cex.axis = 1.5)

# Plot individual lines
for (i in 1:length(cond2)){

	points( x[i]+c(-0.18,-0.06,0.06,0.18), cbind(cond1,cond2,cond3,cond4)[i,], col=c(con,opp,nov,ran),bg=c(conL,oppL,novL,ranL), pch=21 )
	lines( x[i]+c(-0.18,-0.06,0.06,0.18), cbind(cond1,cond2,cond3,cond4)[i,], col=rgb(.5,.5,.5,alpha=.5) )
	
}

# Calculate and present means
means <- colMeans(cbind(cond1,cond2,cond3,cond4))

lines( c(-.192,-.168), rep(means[1],2), lwd=3.5, col = dark, lty = 1)
lines( c(-.072,-.048), rep(means[2],2), lwd=3.5, col = dark, lty = 1)
lines( c(.048,.072), rep(means[3],2), lwd=3.5, col = dark, lty = 1)
lines( c(.168,.192), rep(means[4],2), lwd=3.5,col = dark, lty = 1)

```

<a name="S5"></a>
&nbsp;

#####**5. Block-by-block changes in performance (Moving & Static versions) ** 

```{r}

# specify colours
mycol <- rgb(100, 100, 100, max = 255, alpha = 80)
mcol <- rgb(100, 100, 100, max = 255, alpha = 120)
mcol <- rgb(.1,.1,.1, alpha = 0.7)


### Moving Target Version

dmm <- dm %>% group_by(ID, block, condition) %>% dplyr::summarize(d_m = mean(distance), d_md = median(distance))

dt <- dmm

dt$d_md <- dt$d_md * 100
levels(dt$condition)[levels(dt$condition)=="Incongruent"] <- "Opposite"


# the most general analysis: ANOVA (3 blocks x 4 conditions)
m <- aov_ez(id = "ID", dv = "d_md", within = c("block", "condition"), data = dt)
m
print(summary(m)) # degrees of freedom: levels - 1,  (subjects - 1) * (levels - 1)

# post-hoc comparisons
em <- emmeans(m, ~ block * condition)
print(em)
em_pairs <- pairs(em, adjust = "Bonferroni") # "none"
#print(em_pairs) # very large list - overalps with the analysis below


# conditions assessed separately (blocks-split)
for (c in unique(dt$condition)) {
dc <- dt[dt$condition == c,]

p <- xyplot(d_md ~ block, group = ID, data = dc, ty=c("l", "p"), ylab = "distance to target", ylim = c(0,9.5), main = paste(c), scales=list(x=list(limits=c(0.8,3.2), at = 1:3, alternating=1, cex = 1.3), y=list(cex = 1.3)), panel = function(x, y, ...) {
       panel.xyplot(x, y, ..., alpha = 0.85, lwd = 1.3)
       panel.linejoin(x, y, fun = mean, horizontal = FALSE,
       col= mcol, lty=1, lwd=6)})

# export 4 x 4 cm
print(p)
print(c)

m <- aov_ez(id = "ID", dv = "d_md", within = "block", data = dc, return = "nice")
print(m)
m <- aov_ez(id = "ID", dv = "d_md", within = "block", data = dc)
print(summary(m)) 


#post-hoc comparisons
em <- emmeans(m, ~ block)
print(em)
em_pairs <- pairs(em, adjust = "Bonferroni") 
print(em_pairs)
}


### Static Target Version

dss <- ds %>%
    
    group_by(ID, block, trial) %>%  
    arrange(desc(timestamp)) %>% 
    slice(1)

dss <- dss %>% group_by(ID, block, condition) %>% dplyr::summarize(h_m = mean(target_num), h_md = median(target_num)) # md i m the same

dt <- dss
levels(dt$condition)[levels(dt$condition)=="Incongruent"] <- "Opposite"

# the most general analysis: ANOVA (3 blocks x 4 conditions)
m <- aov_ez(id = "ID", dv = "h_m", within = c("block", "condition"), data = dt)
m
print(summary(m)) # degrees of freedom: levels - 1,  (subjects - 1) * (levels - 1)

# post-hoc comparisons
em <- emmeans(m, ~ block)
print(em)
em_pairs <- pairs(em, adjust = "Bonferroni") 
# print(em_pairs) # very large list - overalps with the analysis below


# conditions assessed separately (blocks-split)
for (c in unique(dt$condition)) {
dc <- dt[dt$condition == c,]

p <- xyplot(h_m ~ block, group = ID, data = dc, ty=c("l", "p"),  ylab = "hits on target", ylim = c(0,20), main = paste(c), scales=list(x=list(limits=c(0.8,3.2), at = 1:3, alternating=1, cex = 1.3), y=list(cex = 1.3)), panel = function(x, y, ...) {
       panel.xyplot(x, y, ..., alpha = 0.85, lwd = 1.3)
       panel.linejoin(x, y, fun = mean, horizontal = FALSE,
       col= mcol, lty=1, lwd=6)})

print(p)
print(c)

m <- aov_ez(id = "ID", dv = "h_m", within = "block", data = dc, return = "nice")
print(m)
m <- aov_ez(id = "ID", dv = "h_m", within = "block", data = dc)
print(summary(m)) 

#post-hoc comparisons
em <- emmeans(m, ~ block)
print(em)
em_pairs <- pairs(em, adjust = "Bonferroni") 
print(em_pairs)
}

```

<a name="S6"></a>
&nbsp;

#####**6. Relationship between performance in the Moving and Static Target versions** 


```{r}

# get chance level line
maxdev <- max(dm$distance)
chance_level = maxdev/2

# prepare Moving Target data
dmm <- dm %>% group_by(ID, block, condition) %>% dplyr::summarize(d_m = mean(distance), d_md = median(distance))

# prepare Static Target data
dss <- ds %>%
    group_by(ID, block, trial) %>%  
    arrange(desc(timestamp)) %>% 
    slice(1)
dss <- dss %>% group_by(ID, block, condition) %>% dplyr::summarize(hits = mean(target_num))

# merge data from both versions
dd <- merge(dmm, dss, by = c("ID", "block", "condition"))

# basic exploration - subjects x blocks (optional)
dt <- dd
for (c in unique(dt$condition)) {
dc <- dt[dt$condition == c,]
#print(c)
#print(shapiro.test(dc$d_md))
#print(shapiro.test(dc$hits))
#print(cor.test(dc$d_md, dc$hits, method = "spearman"))
#scatter.smooth(dc$d_md, dc$hits, xlim = c(0.0, 0.1), ylim = c(0, 15),main = paste(c), xlab = "distance to target", ylab = "hits")
}


### Subject level data
dmS <- dmm %>% group_by(ID, condition) %>% dplyr::summarize(d_m = mean(d_m), d_md = mean(d_md))
dsS <- dss %>% group_by(ID, condition) %>% dplyr::summarize(hits = mean(hits))
DD <- merge(dmS, dsS, by = c("ID", "condition"))

scatter.smooth(DD$d_md, DD$hits, xlim = c(0.0, 0.08), ylim = c(0, 15), xlab = "distance to target", ylab = "hits")

dt <- DD
dt$d_md <- dt$d_md * 100

## analyze relationship
for (c in unique(dt$condition)) {
dc <- dt[dt$condition == c,]
print(c)
print(shapiro.test(dc$d_md))
print(shapiro.test(dc$hits))
print(cor.test(dc$d_md, dc$hits, method = "spearman"))
#scatter.smooth(dc$d_md, dc$hits, xlim = c(0.0, 0.1), ylim = c(0, 15),main = paste(c), xlab = "distance to target", ylab = "hits")
}


### Visualise data
sp <- ggscatter(dt, x = "d_md", y = "hits",
  ellipse = T, ellipse.level = 0.9, ellipse.type = "t", # norm/euclid #level095
  ellipse.alpha = 0.2, ellipse.border.remove = T,
  #add = "reg.line",             # Add regression line
  #conf.int = T,                # Add confidence interval
  color = "condition", size = c(2.2), alpha = 0.7, # Color by groups
  palette = c(con,opp,nov,ran), 
  shape = "condition") + scale_shape_manual(values=c(19, 15, 17, 23)) 
sp

# export: 4.1 x 4.6 (1.8 point size)
# export: 4.7 x 5.3 (2.2 point size)

```

<a name="S7"></a>
&nbsp;

#####**7. Relationship between performance and speed of actions** 

```{r}

### Moving Target - prepare speed of action data
dmm_speed <- dm %>%
    group_by(ID, block, trial) %>%  
    arrange(desc(timestamp)) %>% 
    slice(1)
dmm_speed <- select(dmm_speed, ID, condition, speed_average)
dmm_speed <- dmm_speed %>% group_by(ID, condition) %>% dplyr::summarize(speed = mean(speed_average))


dmm <- dm %>% group_by(ID, condition) %>% dplyr::summarize(d_md = median(distance))
dmC <- merge(dmm, dmm_speed, by = c("ID", "condition"))
dm_s <- dmC # speed data

## differences in speed of actions between sensorimotor conditions
levels(dm_s$condition)[levels(dm_s$condition)=="Incongruent"] <- "Opposite"
m <- aov_ez(id = "ID", dv = "speed", within = "condition", data = dm_s)
m
print(summary(m)) # degrees of freedom: levels - 1,  (subjects - 1) * (levels - 1)

#post-hoc comparisons
em <- emmeans(m, ~ condition)
print(em)
em_pairs <- pairs(em, adjust = "Bonferroni") 
print(em_pairs)

# visualise results
xyplot(speed ~ condition, group = ID, data = dm_s, ylab = "Speed of manual action", ty=c("l", "p"), ylim = c(0,2.3), panel = function(x, y, ...) {
       panel.xyplot(x, y, ..., alpha = 0.7, lwd = 1.3)
       panel.linejoin(x, y, fun = mean, horizontal = FALSE,
       col= mcol, lty=1, lwd=6)})


# speed & performance at different sensorimotor conditions
par(mar=c(2.1,2.1,1.1,1.1),mfrow=c(2,2))
dt <- dm_s 
dt$d_md <- dt$d_md *100

for (c in unique(dt$condition)) {
dc <- dt[dt$condition == c,]

print(c)

print(shapiro.test(dc$speed))
test <- shapiro.test(dc$speed)
norm1 <- (test$p.value > 0.05)

print(shapiro.test(dc$d_md))
test <- shapiro.test(dc$d_md)
norm2 <- (test$p.value > 0.05)

ifelse((norm1 == FALSE | norm2 == FALSE), R_method <- "spearman", R_method <- "pearson")

cor <- cor.test(dc$speed, dc$d_md, method = R_method, exact = F)
print(cor)
p_val <- round(cor$p.value, 3)
p <- scatter.smooth(dc$speed, dc$d_md, ylim = c(0,8), main = paste(c, " p =", p_val), xlab = "Speed", ylab = "Distance to target")
p
}

### Static Target - prepare speed of action data
dss_speed <- ds %>%
    group_by(ID, block, trial) %>%  
    arrange(desc(timestamp)) %>% 
    slice(1)
dss_speed <- select(dss_speed, ID, condition, speed_average)
dss_speed <- dss_speed %>% group_by(ID, condition) %>% dplyr::summarize(speed = mean(speed_average))

dss <- ds %>%
    group_by(ID, block, trial) %>%  
    arrange(desc(timestamp)) %>% 
    slice(1)

dss <- dss %>% group_by(ID, condition) %>% dplyr::summarize(h_m = mean(target_num), h_md = median(target_num))

dsC <- merge(dss, dss_speed, by = c("ID", "condition"))

# speed at different conditions (general)
ds_s <- dsC %>% group_by(ID, condition) %>% dplyr::summarise(hits = mean(h_m), speed = mean(speed))

## differences in speed of actions between sensorimotor conditions
levels(ds_s$condition)[levels(ds_s$condition)=="Incongruent"] <- "Opposite"
m <- aov_ez(id = "ID", dv = "speed", within = "condition", data = ds_s)
m
print(summary(m)) # degrees of freedom: levels - 1,  (subjects - 1) * (levels - 1)

#post-hoc comparisons
em <- emmeans(m, ~ condition)
print(em)
em_pairs <- pairs(em, adjust = "Bonferroni") 
print(em_pairs)

# visualise results
xyplot(speed ~ condition, group = ID, data = ds_s, ylab = "speed of manual action", ty=c("l", "p"), ylim = c(0,2.3), panel = function(x, y, ...) {
       panel.xyplot(x, y, ..., alpha = 0.7, lwd = 1.3)
       panel.linejoin(x, y, fun = mean, horizontal = FALSE,
       col= mcol, lty=1, lwd=6)})



# speed & performance at different sensorimotor conditions
dt <- ds_s 
par(mar=c(2.1,2.1,1.1,1.1),mfrow=c(2,2))

for (c in unique(dt$condition)) {
dc <- dt[dt$condition == c,]

print(c)

print(shapiro.test(dc$speed))
test <- shapiro.test(dc$speed)
norm1 <- (test$p.value > 0.05)

print(shapiro.test(dc$hits))
test <- shapiro.test(dc$hits)
norm2 <- (test$p.value > 0.05)

ifelse((norm1 == FALSE | norm2 == FALSE), R_method <- "spearman", R_method <- "pearson")
cor <- cor.test(dc$speed, dc$hits, method = R_method, exact = F)
print(cor)
p_val <- round(cor$p.value, 3)

p <- scatter.smooth(dc$speed, dc$hits, ylim = c(0,18), main = paste(c, " p =", p_val), xlab = "Speed", ylab = "Hits on target")
p
}

# general comparisons of speed at different tasks
dss_ID_speed <- dss_speed %>% group_by(ID) %>% dplyr::summarize(s_speed = mean(speed))
dmm_ID_speed <- dmm_speed %>% group_by(ID) %>% dplyr::summarize(m_speed = mean(speed))
ID_speed <- merge(dss_ID_speed, dmm_ID_speed, by = "ID")
mean(ID_speed$s_speed)
mean(ID_speed$m_speed)
t.test(ID_speed$m_speed, ID_speed$s_speed, paired = T)

```

<a name="S8"></a>
&nbsp;

#####**8. Merging CFS and Sensorimotor Mastery data** 

```{r}

# general
dmm <- dm %>% group_by(ID, block, condition) %>% dplyr::summarize(dist = median(distance))
dt <- dmm

### STATIC - target num
dss <- ds %>%
    group_by(ID, block, trial) %>%  
    arrange(desc(timestamp)) %>% 
    slice(1)

dss <- dss %>% group_by(ID, block, condition) %>% dplyr::summarize(hits = mean(target_num))

ddd <- dd  %>% group_by(ID, condition) %>% dplyr::summarize(d_md = mean(d_md), hits = mean(hits))

# ID level
tot <- merge(cfs_ID, ddd, by = c("ID", "condition"))
colnames(tot)[6] <- "dist" # distance from Moving Target version
#write.table(tot, file = "smcvr_cfs_sm_data_ID.csv" )

# blocks level
tot_blocks <- merge(cfs, dd, by = c("ID", "condition", "block"))
colnames(tot_blocks)[8] <- "dist" # distance from Moving Target version
tot_blocks <- select(tot_blocks, -c("d_m"))
#write.table(tot_blocks, file = "smcvr_cfs_sm_data_blocks.csv" )


```


