---
title: <font size="5">**SMCVR -- Continuous Flash Suppression & Sensorimotor Mastery - associations**</font> 
author: <br> <font size="4"> Pawel Motyka (Polish Academy of Sciences) </font> <br>  *pmotyka@psych.pan.pl* 
date: <font size="3"> September 2022  </font>
output: html_document
chunk_output_type: console
editor_options: 
  chunk_output_type: console
--- 
&nbsp;
<font size="4">
**List of sections**:

1. Load the required packages and data [S1](#S1)
2. Derive the single sensorimotor performance index [S2](#S2)
3. Assess associations between sensorimotor performance and detection times - separate conditions [S3](#S3)
4. Assess associations between sensorimotor performance and detection times - separate conditions and blocks [S4](#S4)
5. Focus on the novel condition specifically [S5](#S5)


<a name="S1"></a>
&nbsp;

#####**1. Load the required packages and data** 

```{r, message = FALSE, warning = FALSE}

# load packages
library(here, warn.conflicts = FALSE, quietly=TRUE)
library(dplyr, warn.conflicts = FALSE, quietly=TRUE)
library(lattice, warn.conflicts = FALSE, quietly=TRUE)
library(afex, warn.conflicts = FALSE, quietly=TRUE)
library(emmeans, warn.conflicts = FALSE, quietly=TRUE)
library(ggpubr, warn.conflicts = FALSE, quietly=TRUE)
#library(tidyr, warn.conflicts = FALSE, quietly=TRUE)

# set working directory
data_dir <- paste0(here(),"/_data")
setwd(data_dir)

# read preprocessed data 
dt <- read.table(file = "smcvr_cfs_sm_data_blocks.csv" ) 
dtS <- read.table(file = "smcvr_cfs_sm_data_ID.csv" )

# recode variables
levels(dt$condition)[levels(dt$condition)=="Incongruent"] <- "Opposite"
levels(dtS$condition)[levels(dtS$condition)=="Incongruent"] <- "Opposite"
dt$dist <- dt$dist * 100 # to cm
dtS$dist <- dtS$dist * 100 # to cm


```

<a name="S2"></a>
&nbsp;

#####**2. Derive the single sensorimotor perfromance index** 


```{r, warning = FALSE}

# According to the preregistration: "If performance measures from both tasks turn out to be highly correlated, a single index (e.g., the average of ranks or standardised values from both tasks) will be determined and used in exploratory analyses".

### Subjects-level

## get average of ranks from Moving and Static Target versions

# Distance rank (Moving Target version)
dtS$dist_rank <- rank(-dtS$dist)
scatter.smooth(dtS$dist, dtS$dist_rank)

# Hits rank (Static Target version)
dtS$hits_rank <- rank(dtS$hits)
scatter.smooth(dtS$hits, dtS$hits_rank)

# associations between ranks
for (c in unique(dtS$condition)) {
dtSC <- dtS[dtS$condition == c,]

# check correlations between ranks
scatter.smooth(dtSC$dist_rank, dtSC$hits_rank, main = paste0("correlation: ", c))
cor <- cor.test(dtSC$dist_rank, dtSC$hits_rank, method = "spearman")
print(cor) }

# average rank from both tasks
dtS$mastery <- (dtS$dist_rank + dtS$hits_rank) / 2

# normalize index (0-10)
dtS$mastery <- ((dtS$mastery - min(dtS$mastery)) / (max(dtS$mastery) - min(dtS$mastery))) * 10
hist(dtS$mastery, breaks = 20, main = "Sensorimotor Mastery index (subjects-level)")



### Blocks-level 

## get average of ranks from Moving and Static Target versions

# Distance rank (Moving Target version)
dt$dist_rank <- rank(-dt$dist)
scatter.smooth(dt$dist, dt$dist_rank)

# Hits rank (Static Target version)
dt$hits_rank <- rank(dt$hits)
scatter.smooth(dt$hits, dt$hits_rank)

# check correlations between ranks
par(mfrow=c(2,3))

for (c in unique(dt$condition)) {
  
dC <- dt[dt$condition == c,]

for (b in unique(dC$block)) {
  
dCb <- dC[dC$block == b,]
  
print(paste0("condition: ", c, " block: ", b))
cor <- cor.test(dCb$dist_rank, dCb$hits_rank, method = "spearman")
p_val <- round(cor$p.value, 3)
scatter.smooth(dCb$dist_rank, dCb$hits_rank, main = paste0(c, " block:", b, ", p = ", p_val))
print(cor)
}
}

rm(dCb)
rm(dC)
rm(cor)


### create single index

# average rank from both tasks
dt$mastery <- (dt$dist_rank + dt$hits_rank) / 2

# normalize ranks (0-10)
dt$mastery <- ((dt$mastery - min(dt$mastery)) / (max(dt$mastery) - min(dt$mastery))) * 10
par(mfrow=c(1,1))
hist(dt$mastery, breaks = 20, main = "Sensorimotor Mastery index (blocks-level)")


```



<a name="S3"></a>
&nbsp;

#####**3. Assess associations between sensorimotor performance and detection times - separate conditions** 

```{r, warning = FALSE}

par(mar=c(2.1,2.1,1.1,1.1),mfrow=c(2,2))

### General - Sensorimotor Mastery Index

## Normalized detection times

for (c in unique(dtS$condition)) {

dc <- dtS[dtS$condition == c,]

print(c)

print(shapiro.test(dc$time_n))
test <- shapiro.test(dc$time_n)
norm1 <- (test$p.value > 0.05)

print(shapiro.test(dc$mastery))
test <- shapiro.test(dc$mastery)
norm2 <- (test$p.value > 0.05)

ifelse((norm1 == FALSE | norm2 == FALSE), R_method <- "spearman", R_method <- "pearson")

cor <- cor.test(dc$time_n, dc$mastery, method = R_method)
p_val <- round(cor$p.value, 3)
print(cor)

p <- scatter.smooth(dc$mastery,dc$time_n, xlim = c(0,10), main = paste(c, ", p = ", p_val), xlab = "mastery", ylab = "breakthrough time (normalized)")
p
}


## Raw detection times (additional exploration)

for (c in unique(dtS$condition)) {

dc <- dtS[dtS$condition == c,]

print(c)

print(shapiro.test(dc$time))
test <- shapiro.test(dc$time)
norm1 <- (test$p.value > 0.05)

print(shapiro.test(dc$mastery))
test <- shapiro.test(dc$mastery)
norm2 <- (test$p.value > 0.05)

ifelse((norm1 == FALSE | norm2 == FALSE), R_method <- "spearman", R_method <- "pearson")

cor <- cor.test(dc$time, dc$mastery, method = R_method)
p_val <- round(cor$p.value, 3)
print(cor)

p <- scatter.smooth(dc$mastery,dc$time, xlim = c(0,10), main = paste(c, ", p = ", p_val), xlab = "mastery", ylab = "breakthrough time (s)")
p
}

### Optional: check separately for Moving and Static Target Versions (no significant associations)

optional_check = F

if (optional_check == F) {
  
# Normalized detection times - Moving Target Version

for (c in unique(dtS$condition)) {

dc <- dtS[dtS$condition == c,]

print(c)

print(shapiro.test(dc$time_n))
print(shapiro.test(dc$dist))
cor <- cor.test(dc$time_n, dc$dist, method = "spearman")
p_val <- round(cor$p.value, 3)
print(cor)

p <- scatter.smooth(dc$dist,dc$time_n, xlim = c(0,10), main = paste(c, ", p = ", p_val), xlab = "distance", ylab = "breakthrough time (normalized)")
p
}
 
# Normalized detection times - Static Target Version

for (c in unique(dtS$condition)) {

dc <- dtS[dtS$condition == c,]

print(c)

print(shapiro.test(dc$time_n))
print(shapiro.test(dc$hits))
cor <- cor.test(dc$time_n, dc$hits, method = "spearman")
print(cor)
p_val <- round(cor$p.value, 3)
p <- scatter.smooth(dc$hits,dc$time_n, xlim = c(0,20), main = paste(c, ", p = ", p_val), xlab = "hits", ylab = "breakthrough time (normalized)")
p
}
}

```



<a name="S4"></a>
&nbsp;

#####**4. Assess associations between sensorimotor performance and detection times - separate conditions and blocks** 

```{r, warning = FALSE}

par(mfrow=c(2,3))

for (c in unique(dt$condition)) {
dc <- dt[dt$condition == c,]

for (b in 1:3) {
  
dcb <- dc[dc$block == b,]  

print(c)
print(b)

print(shapiro.test(dcb$time_n))
test <- shapiro.test(dcb$time_n)
norm1 <- (test$p.value > 0.05)

print(shapiro.test(dcb$mastery))
test <- shapiro.test(dcb$mastery)
norm2 <- (test$p.value > 0.05)

ifelse((norm1 == FALSE | norm2 == FALSE), R_method <- "spearman", R_method <- "pearson")

cor <- cor.test(dcb$mastery, dcb$time_n, method = R_method)
print(cor)
p_val <- round(cor$p.value, 3)

m <- scatter.smooth(dcb$mastery, dcb$time_n, xlim = c(0,10), ylim = c(-2,2), main = paste(c, "- block:", b, " p =", p_val))
}
}


# Plot results for congruent, opposite, and random conditions
con = rgb(.28,.13,.45, alpha = 0.65)
opp = rgb(.18,.44,.56, alpha = 0.8)
nov = rgb(.136,.553,.425, alpha = 0.8)
ran = rgb(.70,.64,.15, alpha = 0.8)

## Congruent
# prepare data
dc <- dt[dt$condition == "Congruent",]
dc$time_n <- as.numeric(dc$time_n)
dc$mastery <- as.numeric(dc$mastery)
dc$block <- as.factor(dc$block)

# specify colours (darkness)
#nov1 = rgb(.136 * 1.4,.553 * 1.4,.425 * 1.4, alpha = 0.8)
con1 = rgb(.28 * 1.6,.13 * 1.6,.45 * 1.6, alpha = 0.8)
con2 = rgb(.28,.13,.45, alpha = 0.8)
con3 = rgb(.28/2,.13/2,.45/2, alpha = 0.8)

## Plot associations
sp <- ggscatter(dc, x = "mastery", y = "time_n",
  ellipse = F, ellipse.level = 0.01, ellipse.type = "t", # norm/euclid #level095
  ellipse.alpha = 0.2, ellipse.border.remove = T,
  mean.point = T, mean.point.size = 4,star.plot = F,
  add = "reg.line",             # Add regression line
  conf.int = T,                # Add confidence interval
  color = "block", size = c(1.0), # Color by groups
  palette = c(con1,con2,con3), 
  shape = "block") + scale_shape_manual(values=c(19, 15, 17)) 
sp


## Opposite
# prepare data
dc <- dt[dt$condition == "Opposite",]
dc$time_n <- as.numeric(dc$time_n)
dc$mastery <- as.numeric(dc$mastery)
dc$block <- as.factor(dc$block)

# specify colours (darkness)
#nov1 = rgb(.136 * 1.4,.553 * 1.4,.425 * 1.4, alpha = 0.8)
opp1 = rgb(.18 * 1.4,.44 * 1.4,.56 * 1.4, alpha = 0.8)
opp2 = rgb(.18,.44,.56, alpha = 0.8)
opp3 = rgb(.18/2,.44/2,.56/2, alpha = 0.8)

## Plot associations
sp <- ggscatter(dc, x = "mastery", y = "time_n",
  ellipse = F, ellipse.level = 0.01, ellipse.type = "t", # norm/euclid #level095
  ellipse.alpha = 0.2, ellipse.border.remove = T,
  mean.point = T, mean.point.size = 4,star.plot = F,
  add = "reg.line",             # Add regression line
  conf.int = T,                # Add confidence interval
  color = "block", size = c(1.0), # Color by groups
  palette = c(opp1,opp2,opp3), 
  shape = "block") + scale_shape_manual(values=c(19, 15, 17)) 
sp


## Random
# prepare data
dc <- dt[dt$condition == "Random",]
dc$time_n <- as.numeric(dc$time_n)
dc$mastery <- as.numeric(dc$mastery)
dc$block <- as.factor(dc$block)

# specify colours (darkness)
#nov1 = rgb(.136 * 1.4,.553 * 1.4,.425 * 1.4, alpha = 0.8)
ran1 = rgb(.70 * 1.1,.64 * 1.1,.15 * 1.1, alpha = 0.8)
ran2 = rgb(.70/1.2,.64/1.2,.15/1.2, alpha = 0.8)
ran3 = rgb(.70/2,.64/2,.15/2, alpha = 0.8)

## Plot associations
sp <- ggscatter(dc, x = "mastery", y = "time_n",
  ellipse = F, ellipse.level = 0.01, ellipse.type = "t", # norm/euclid #level095
  ellipse.alpha = 0.2, ellipse.border.remove = T,
  mean.point = T, mean.point.size = 4,star.plot = F,
  add = "reg.line",             # Add regression line
  conf.int = T,                # Add confidence interval
  color = "block", size = c(1.0), # Color by groups
  palette = c(ran1,ran2,ran3), 
  shape = "block") + scale_shape_manual(values=c(19, 15, 17)) 
sp



```


<a name="S5"></a>
&nbsp;

#####**5. Focus on the novel condition specifically ** 

```{r, warning = FALSE}

# prepare data
dc <- dt[dt$condition == "Novel",]
dc$time_n <- as.numeric(dc$time_n)
dc$mastery <- as.numeric(dc$mastery)
dc$block <- as.factor(dc$block)

# specify colours (darkness)
#nov1 = rgb(.136 * 1.4,.553 * 1.4,.425 * 1.4, alpha = 0.8)
nov1 = rgb(.136 * 1.25,.553 * 1.25,.425 * 1.25, alpha = 0.8)
nov2 = rgb(.136/1.2,.553/1.2,.425/1.2, alpha = 0.8)
nov3 = rgb(.136/2,.553/2,.425/2, alpha = 0.8)


## Plot associations
sp <- ggscatter(dc, x = "mastery", y = "time_n",
  ellipse = F, ellipse.level = 0.01, ellipse.type = "t", # norm/euclid#level095
  ellipse.alpha = 0.2, ellipse.border.remove = T,
  mean.point = T, mean.point.size = 4.5,star.plot = F,
  add = "reg.line",             # Add regression line
  conf.int = T,                # Add confidence interval
  color = "block", size = c(1.5), # Color by groups
  palette = c(nov1,nov2,nov3), 
  shape = "block") + scale_shape_manual(values=c(19, 15, 17)) 
sp

# for Fig4e change point sizes: 5 & 1.3 (export 4.7 x 5.5)

## check how many people had a better score than in the random condition
d <- dtS

d <- dplyr::select(d, ID, condition, mastery)
d <- tidyr::pivot_wider(d, id_cols = ID, names_from = condition, values_from =  mastery)

d$learned <- (d$Novel > d$Random)
length(d$ID[d$learned == T])


## check the correlation between change in sensorimotor performance (blocks 1-3) and change in detection times (blocks 1-3)

par(mfrow=c(1,1))

# prepare data
d <- dt[dt$block != 2,]
d <- d[d$condition == "Novel",]
d <- select(d, ID, block, condition, time, time_n, mastery)
d <- tidyr::pivot_wider(d, id_cols = ID, names_from = block, values_from = c(mastery,time,time_n))

## mastery: change between the last and the block
d$mastery_change <- d$mastery_3 - d$mastery_1
#hist(d$mastery_change)

## detection times (s): change between the last and first block
d$time_change <- d$time_3 - d$time_1
#hist(d$time_change)

## detection times (normalised): change between the last and first block
d$timeN_change <- d$time_n_3 - d$time_n_1
#hist(d$timeN_change)

# test normality of variables
shapiro.test(d$mastery_change)
shapiro.test(d$time_change)
shapiro.test(d$timeN_change)

# change in mastery vs change in detection times (s)
cor.test(d$mastery_change, d$time_change, method = "spearman")
#scatter.smooth(d$mastery_change, d$time_change)

# change in mastery vs change in detection times (normalised)
cor.test(d$mastery_change, d$timeN_change, method = "pearson")
#scatter.smooth(d$mastery_change, d$timeN_change)

```






